---
title: "Untitled"
author: "Dan Aurell"
date: "2023-03-23"
output: 
  html_document:
    keep_md: yes
---


```{r}
library(tidyverse)
```



```{r}
datum <- read.csv("../Assign_4_data_wrangling/data/Microbiome.csv")

str(datum)
```


Q1. Select the following columns.

```{r eval=FALSE}
datum %>% 
  select(c(OTU, SampleID, Abundance, Crop, Compartment, DateSampled, GrowthStage, Treatment, Rep, Fungicide, Kingdom, Phylum, Class, Order, Family, Genus, Species, Taxonomy))
```



Q2. Calculate the mean percent relative abundance across all OTUs by compartment using the pipe operator.

```{r}
datum %>% 
  group_by(Compartment) %>% 
  summarize(Abundance = mean(Abundance))
```


Q3. Calculate the mean percent relative abundance across all OTUs by compartment and fungicide using the pipe operator.

```{r}
datum %>% 
  group_by(Compartment, Fungicide) %>% 
  summarize(Abundance = 100*mean(Abundance))
```

Q4. Restrict the previous analysis to just the fungal class Dothideomycetes.

```{r}
datum %>% 
  filter(Class == "Dothideomycetes") %>% 
  group_by(Compartment, Fungicide) %>% 
  summarize(Abundance = 100*mean(Abundance))
```

Q5. Now do the same analysis but for Orders in the class Dothideomycetes.

```{r}
datum %>% 
  filter(Class == "Dothideomycetes") %>% 
  group_by(Order, Compartment, Fungicide) %>% 
  summarize(Abundance.m = 100*mean(Abundance))
```

Q6. Calculate the standard error for each mean in Q5.

```{r}
df <- datum %>% 
  filter(Class == "Dothideomycetes") %>% 
  group_by(Order, Compartment, Fungicide) %>% 
  summarize(Abundance.m = 100*mean(Abundance), 
            n=n(),
            se = sd(Abundance)/sqrt(n)
            )
```


Q7. Select the columns Order, Compartment, Fungicide and Mean from the output of Q6 and Pivot the dataset to wide format so you have Compartment as column headers

```{r}
df <- df %>% 
  select(Order, Compartment, Fungicide, Abundance.m)

df %>% 
  pivot_wider(names_from = c(Compartment), values_from = Abundance.m)
```


Q8. Recreate this plot using your dataset in Q7. 

```{r}
order <- datum %>% 
  filter(Class == "Dothideomycetes") %>%
  group_by(Order) %>% 
  summarize(Abundance.m = 100*mean(Abundance)) %>% 
  arrange(Abundance.m) %>% 
  select(Order)

order <- order$Order

plotdatum <- datum %>% 
  filter(Class == "Dothideomycetes") %>%
  group_by(Order, Fungicide) %>% 
  summarize(Abundance.m = 100*mean(Abundance))

plotdatum$Order <- factor(plotdatum$Order, levels=order)
  
```


```{r}
plotdatum %>% 
  ggplot(aes(x=Order, y = Abundance.m, fill = Fungicide)) +
  geom_col(position = "dodge") +
  xlab(NULL) +
  ylab("Percent relative abundance") +
  coord_flip()
```