---
title: "Basic Analysis"
author: "Dan Aurell"
date: "2023-04-05"
output:   
  html_document: 
    keep_md: yes
---

# Question 1

*Background*

* Varroa mites are bad for honey bee colonies. 
* To determine whether Varroa mites are at damaging levels, we take a sample of 1/2 cup of bees and count the number of mites in those samples.
* Several treatments are used in the industry:
  * Apivar (amitraz-based, registered, acts slowly)
  * Amitraz E.C. (amitraz-based, not registered for use in bees, reputed to act quickly and effectively)
  * Apiguard (thymol-based, registered, acts quickly)

*Data*

* For simplicity, I'm going to look at just one response variable for this assignment: The number of Varroa mites per sample of ~300 bees.


*Tested treatments*

* Apivar (Standard treatment)
* Combination of Apivar & Apiguard (Test treatment)
* Amitraz E.C. (Positive control, this is the one beekeepers say works the best)

*Experimental design*

* Colonies in the same apiary assigned to receive the different treatments
* Repeated-measures design
* Varroa infestation rate measured for each colony on Day 0 (Day of treatment), on Day 21, and Day 42

*Objectives*

* Determine whether a Combination treatment can control Varroa more effectively than Apivar on its own
* Determine whether a Combination treatment can control Varroa as effectively as Amitraz E.C.
* Determine whether Apivar is more or less effective against Varroa than Amitraz E.C.

*Hypotheses*

* A Combination treatment can control Varroa more effectively than Apivar on its own
* A Combination treatment can control Varroa as effectively as Amitraz E.C.
* Apivar is less effective than Amitraz E.C.

*Statistical plan to test hypotheses*

* Consider experiment day and treatment as categorical explanatory variables (allows for nonlinear response of Varroa through time)
* I previously found evidence of overdispersion when modeling the data as Poisson, so we'll use a negative binomial model instead
* Since it's a repeated measures design (each colony measured thrice), we'll include a random effect of Colony ID
* Since we have a pre-post design, we're not looking for a treatment effect, statistically speaking (this would mean that a difference between treatments existed on Day 0 and all other days). What we're actually asking is... is there an interaction between Day and Treatment?
* We'll report an overall ANOVA P-value for the Day-Treatment interaction
* Since there was a significant Day-Treatment interaction, we will use emmeans() to compare the treatments within each day.
* Using Emmeans, we will do a pairwise contrast (post-hoc tes) within each day; therefore, three comparisons are made within Day 0 (and for Day 21 and 42); therefore, there will be multiplicity adjustment for three comparisons. Emmeans by default uses the Tukey's Honestly Significant Differences adjustment.

# Question 2
See code below


```{r}
library(tidyverse)
library(lme4)
library(emmeans)
```

Read data
```{r}
datum <- read.csv("datum_simplified.csv")
```

Set variables as factors
```{r}
datum$trt <- factor(datum$trt , levels=c("Apivar", "AEC", "Apiguard_Api"))
datum$days <- factor(datum$days)
```


Fit model and null model
```{r}
tempdatum <- datum %>% filter(
  days != "10"
) # Because we did not measure mite infestation of adult bees on Day 10

# "model of phoretic mite number per sample fit with negative binomial family to account for count data with overdispersion"
m.phor.nb1 <- glmer.nb(mites_recov_total ~ 
                     trt + 
                     days + 
                     trt:days + 
                     (1 | col_no), 
                   data = tempdatum)
summary(m.phor.nb1)

isREML(m.phor.nb1) # glmer.nb models are always fit by log-likelihood
```

Test significance of interaction
```{r}
# Fit model that omits trt:days interaction
m.phor.nb0 <- glmer.nb(mites_recov_total ~ 
                     trt + 
                     days + 
                     # trt:days + 
                     (1 | col_no), 
                   data = tempdatum)
summary(m.phor.nb0)

anova(m.phor.nb0, m.phor.nb1) # There was a significant interaction between day and treatment (p=0.0001481) when tested by ANOVA. Therefore we analyzed each date separately.
```


Emmeans saved to an object to send to ggplot for plotting
```{r}
emplot_phor <- emmip(m.phor.nb1, trt~days, type = "response", CIs = TRUE, plotit = FALSE)
emplot_phor$days <- as.numeric(as.character(emplot_phor$days))
emplot_phor$trt <- factor(emplot_phor$trt, levels = c("Apivar", "AEC", "Apiguard_Api"))
levels(emplot_phor$trt) # check order of factors
```

Make plot
```{r, echo=FALSE}
pd <- position_dodge(2)

p <- emplot_phor %>% 
ggplot(mapping = aes(y = yvar/3, x = days, color = trt)) +  
  # , group = trt probably not necessary
  scale_colour_manual(
    limits = c("Apivar", "AEC", "Apiguard_Api"),
    labels = c("Apivar", "Amitraz E.C.", "Combination"),
    values = c("black", "black", "#00BA38"), 
    aesthetics = "color"
  ) +

  scale_linetype_manual(
    limits = c("Apivar", "AEC", "Apiguard_Api"),
    labels = c("Apivar", "Amitraz E.C.", "Combination"),
    values=c("solid", "dashed", "solid")
    ) +
  
  scale_shape_manual(
    limits = c("Apivar", "AEC", "Apiguard_Api"),
    labels = c("Apivar", "Amitraz E.C.", "Combination"),
    values = c(16, 21, 16)) +  
  
  geom_hline(yintercept = 3, linetype = 2, colour = "red") +
  geom_line(aes(linetype = trt), position=pd, size = 1.5) +
  geom_errorbar(aes(ymin=LCL/3, ymax=UCL/3), width=2, position=pd, size = 1) +
  geom_point(aes(shape=trt), position=pd, fill="white", size = 3, stroke = 1) +
  scale_x_continuous(breaks = c(0, 21, 42)) +
  theme_classic(base_size = 30) +
  ylab("Varroa / 100 bees") +
  xlab("Experiment day")


p + theme(legend.title = element_blank()) # No need for "treatment" label
ggsave("p.phor_legend.png", width = 12, height = 8.625, units = "in")

p + theme(legend.position = "none") # For when I want to export a version sans legend
ggsave("p.phor_no_legend.png", width = 12, height = 8.625, units = "in")

```

Emmeans to make statistical comparisons between treatments within day
P-values and estimates of effect size ("times as many Varroa per sample") come from here.
```{r}
emmeans(m.phor.nb1, pairwise ~ trt | days, type = "response")$contrasts
confint(emmeans(m.phor.nb1, pairwise ~ trt | days, type = "response"))$contrasts
```

# Question 3

Based on the output above, I would make interpretations like the following:

* On Day-0, we did not observe significant differences between any treatment groups.
* On Day-21, Apivar colonies had 5.1 (2.1-12.4; 95%C.I.) times as many Varroa mites as Amitraz E.C. colonies (p<0.001).

