---
title: "Basic Analysis"
author: "Dan Aurell"
date: "2023-04-05"
output:   
  html_document: 
    keep_md: yes
---

```{r}
library(tidyverse)
library(lme4)
library(emmeans)
```

Read data
```{r}
datum <- read.csv("datum_simplified.csv")
```

Set variables as factors
```{r}
datum$trt <- factor(datum$trt , levels=c("Apivar", "AEC", "Apiguard_Api"))
datum$days <- factor(datum$days)
```


Since pre-treatment values will be used as covariates, make column of Day-0 values

NOT SURE IF THIS IS ACTUALLY USED
Plus, it gives the models VERY different behavior if it's used or not
```{r}
# Method:
# Extract row based on colony number of row i and day = 0
# Write the value of the column of interest (say bees to start with) into a pre_bees column

#    phor
for(i in 1:nrow(datum)) {
  datum$pre_phor[i] <- datum[which(datum$days == 0 & datum$col_no == datum$col_no[i]), ]$phor
}

#    infest_200
for(i in 1:nrow(datum)) {
  datum$pre_infest_200[i] <- datum[which(datum$days == 0 & datum$col_no == datum$col_no[i]), ]$infest_200
}

#    mites_tot
for(i in 1:nrow(datum)) {
  datum$pre_mites_tot[i] <- datum[which(datum$days == 0 & datum$col_no == datum$col_no[i]), ]$mites_tot
}

#    bees_frames
for(i in 1:nrow(datum)) {
  datum$pre_bees_frames[i] <- datum[which(datum$days == 0 & datum$col_no == datum$col_no[i]), ]$bees_frames
}


```

Fit model and null model
```{r}
tempdatum <- datum %>% filter(
  days != "10"
) # Because we did not measure mite infestation of adult bees on Day 10

# "model of phoretic mite number per sample fit with negative binomial family to account for count data with overdispersion"
m.phor.nb1 <- glmer.nb(mites_recov_total ~ 
                     trt + 
                     days + 
                     trt:days + 
                     # pre_wk_frames + 
                     # pre_bees_frames + 
                     (1 | col_no), 
                   data = tempdatum)
summary(m.phor.nb1)

isREML(m.phor.nb1) # glmer.nb models are always fit by log-likelihood
```

Test significance of interaction
```{r}
# Fit model that omits trt:days interaction
m.phor.nb0 <- glmer.nb(mites_recov_total ~ 
                     trt + 
                     days + 
                     # trt:days + 
                     # pre_wk_frames + 
                     # pre_bees_frames + 
                     (1 | col_no), 
                   data = tempdatum)
summary(m.phor.nb0)

anova(m.phor.nb0, m.phor.nb1) # There was a significant interaction between day and treatment (p=0.0001481) when tested by ANOVA. Therefore we analyzed each date separately.
```

Include this since I did originally
```{r}
m.phor.nb2 <- glmer.nb(mites_recov_total ~ 
                     trt + 
                     days + 
                     trt:days + 
                     # pre_wk_frames + 
                     pre_bees_frames + 
                     (1 | col_no), 
                   data = tempdatum)
summary(m.phor.nb2)
```


Emmeans saved to an object to send to ggplot for plotting
```{r}
emplot_phor <- emmip(m.phor.nb1, trt~days, type = "response", CIs = TRUE, plotit = FALSE)
emplot_phor$days <- as.numeric(as.character(emplot_phor$days))
emplot_phor$trt <- factor(emplot_phor$trt, levels = c("Apivar", "AEC", "Apiguard_Api"))
levels(emplot_phor$trt) # check order of factors
```

Emmeans to make statistical comparisons between treatments within day
P-values and estimates of effect size ("times as many Varroa per sample") come from here.
```{r}
emmeans(m.phor.nb1, pairwise ~ trt | days, type = "response")$contrasts
confint(emmeans(m.phor.nb1, pairwise ~ trt | days, type = "response"))$contrasts
```

```{r}
emmeans(m.phor.nb2, pairwise ~ trt | days, type = "response")$contrasts
confint(emmeans(m.phor.nb2, pairwise ~ trt | days, type = "response"))$contrasts
```

DAY-0 SUMMARY
On Day-0 we did not observe any significant differences across treatment groups (p>0.98).

DAY-21 SUMMARY
On Day-21, there was a trend toward fewer mites in combination-treated colonies than in Apivar-treated colonies (though not significantly; p = 0.17).

We also found that AEC colonies had significantly fewer mites than either Apivar or combination-treated colonies. 

AEC colonies had 0.20 times as many mites (0.086-0.49; 95%C.L.) as Apivar colonies (p=0.0001)

AEC colonies had 0.35 times as many mites (0.15-0.86; 95%C.L.) as combination-treated colonies (p=0.018).

Combination-treated colonies had 0.58 times as many mites (0.28-1.18; 95%C.L.) as Apivar-treated colonies; however, this result was not statistically significant (p=0.17)

DAY-42 SUMMARY
On Day-42 we found that combination treated colonies had a trend toward lower mite counts than Apivar treated colonies, though this was not statistically significant (p=0.13). Combination treated colonies had significantly lower mite counts than AEC treated colonies (p=0.031). We did not observe a difference between Apivar and AEC treated colonies (p=0.84).

Combination treated colonies had 0.55 times the number of mites (0.27-1.13; 95%C.L.) as Apivar treated colonies, though this was not statistically significant (p=0.13).

Combination treated colonies had 0.47 times the number of mites (0.23-0.95; 95%C.L.) as AEC treated colonies (p=0.031).

Plotting
```{r}

```


