---
title: "R Notebook"
output: html_notebook
---

Deviance = error = residuals

Estimate B0, B1, error

How to fit best fit line
- Average residual = 0
- Minimize sum of squared errors (SSE)    

Yi = (y_i-yHat)*2

SSE+ SSR  = (TSS total sum of squares)

The strength regression has to do with the balance between SSE in relation to TXX 

Assumptions of regression
1. Continuous Y
2. Normally distributed residuals
3. *Linear relationship* *Important*
4. No heteroscedasticity
5. Independent samples (No autocorrelation)


*Formal tests only valuable when data sizes are small
  *With large dataset you're going to get 


-------

T-test is 2 groups


What's an ANOVA?
t-test w more than 2 groups

You get an F-statistic from an ANOVA
- testing the significance of [factor]


Basic idea of interactions:
- The effect of one variable depends on the value of another variable



```{r}
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
```

```{r}
data("mtcars")

mtcars %>% 
ggplot(aes(x=wt, y=mpg)) +
  geom_smooth() +
  geom_point()
```


```{r}
lm1 <- lm(mpg~wt, data = mtcars)
summary(lm1) # p-value 1.29e-10

cor.test(mtcars$wt, mtcars$mpg) # A classical test gives the exact same result

ggplot(lm1, aes(y=.resid, x=.fitted)) +
  geom_point()

plot(lm1)

```





