---
title: "R Notebook"
output: html_notebook
---

# General notes

## Tuesday 2023-02-14
ggpubr is a package worth considering using: publication-ready plots

How to find packages not listed in normal CRAN
- e.g., on repository Bioconductor

```{r}
library(tidyverse)
```



```{r}
getwd()

datum <- read.csv("../general_notes/data/corr.csv",na.strings = "NA")

data("mtcars") # this accesses data that are pre-loaded into R
head(airquality)

data() # lists all of the datasets in the data package
```



```{r}
plot(x = mtcars$wt, y = mtcars$mpg,
     xlab = "Car weight",
     ylab = "Miles per gallon",
     font.lab = 6,
     pch = 20
     )
```

Can customize things like

```{r}
mtcars %>%
  ggplot() +
  geom_smooth(aes(x = wt, y = mpg), method = lm, se = FALSE, color = "black") +
  geom_point(aes(x = wt, y = mpg, color = as.factor(cyl)), size = 2, pch = 21) +
  theme_classic() +
  theme(legend.position = "NULL") +
  xlab("Weight") +
  ylab("Miles per gallon")
  
  
```

```{r}
mtcars %>%
  ggplot() +
  geom_smooth(aes(x = wt, y = mpg), method = lm, se = FALSE, color = "black") +
  geom_point(aes(x = wt, y = mpg, color = wt), size = 2, pch = 16) +
  theme_classic() +
  theme(legend.position = "NULL") +
  xlab("Weight") +
  ylab("Miles per gallon") +
  scale_color_gradient(low="forestgreen", high = "black") +
  scale_y_log10()
```

If you're trying to compare means across groups, you can use many different kinds of plots

## Thursday 2023-02-16

```{r}
library(tidyverse)
```

```{r}
bull.richness <- read.csv("./data/Bull_richness.csv", na.strings = NA)
# setting na.strings lets us specify how NAs are coded
```

```{r}
bull.richness.soy.no.till <- bull.richness[bull.richness$Crop == "Soy" 
                                           & bull.richness$Treatment == "No-till",]
```

There are a few variables that are stored as time... tricky to work with in R

So we'll use GrowthStage as our standin for time

Getting used to showing all the data points associated with
*We'll do overplotting using geom_jitter

```{r}
ggplot(bull.richness.soy.no.till, aes(x = GrowthStage, y = richness, color = Fungicide)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(0.2), alpha = 0.5)
```









## Tuesday 2023-02-21

```{r}
library(tidyverse)
```

```{r}
bull.richness <- read.csv("./data/Bull_richness.csv", na.strings = NA)
# setting na.strings lets us specify how NAs are coded
```


```{r}
bull.richness.soy.no.till <- bull.richness[bull.richness$Crop == "Soy" 
                                           & bull.richness$Treatment == "No-till",]
```

```{r}
ggplot(bull.richness.soy.no.till, aes(x = GrowthStage, y = richness, color = Fungicide)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge(0.2), alpha = 0.5) +
  xlab("") +
  ylab("Bulleribasidiaceae richness")
```

### Use stat_summary geom to calculate error bars

```{r}
ggplot(bull.richness.soy.no.till, aes(x = GrowthStage, y = richness, color = Fungicide)) +
  stat_summary(fun= mean, geom = "bar") +
  stat_summary(fun.data = mean_se, geom = "errorbar") +
  xlab("") +
  ylab("Bulleribasidiaceae richness")

# have to adjust the position of the bars, adjust color and fill

```

```{r}
ggplot(bull.richness.soy.no.till, aes(x = GrowthStage, y = richness, fill = Fungicide)) +
  stat_summary(fun= mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  geom_point(position = position_jitterdodge(0.2), pch = 21) +
  xlab("") +
  ylab("Bulleribasidiaceae richness") +
  theme_classic()

# make the shape 21 for filled and outlined circle

```

```{r}
ggplot(bull.richness.soy.no.till, aes(x = GrowthStage, y = richness, fill = Fungicide)) +
  stat_summary(fun= mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  geom_point(position = position_jitterdodge(0.2), pch = 21) +
  scale_color_manual(values = c("blue", "green")) +
  scale_fill_manual(values = c("blue", "green")) +
  xlab("") +
  ylab("Bulleribasidiaceae richness") +
  theme_classic()

```
Maybe we want to use connected lines to represent that it's a time series
```{r}
ggplot(bull.richness.soy.no.till, aes(x = GrowthStage, y = richness, group = Fungicide, color = Fungicide)) +
  stat_summary(fun= mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  geom_jitter(pch=21, width = 0.2) +
  xlab("") +
  ylab("Bulleribasidiaceae richness") +
  theme_classic()
```

Facetting
This was all soybean and no-till. We also have corn and other tillage treatments.

The default is for the scales to be shared across all facets



Allowing scales to be free
```{r}
ggplot(bull.richness, aes(x = GrowthStage, y = richness, group = Fungicide, color = Fungicide)) +
  stat_summary(fun= mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  geom_jitter(pch=21, width = 0.2) +
  xlab("") +
  ylab("Bulleribasidiaceae\n richness") +
  theme_classic() +
  facet_wrap(~Crop*Treatment, scales = "free") 

```

Can set a palette (including for the color blind)
```{r}
cbpalette <- c("black", "green")

bull.richness %>% 
  filter(Crop == "Soy") %>% 
  ggplot(aes(x = GrowthStage, y = richness, group = Fungicide, color = Fungicide)) +
  stat_summary(fun= mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  geom_jitter(pch=21, width = 0.2) +
  scale_color_manual(values =cbpalette) +
  xlab("") +
  ylab("Bulleribasidiaceae\n richness") +
  theme_classic() +
  facet_wrap(~Crop*Treatment, scales = "free") 

```






