---
title: "R Notebook"
output: html_notebook
---

Deviance = error = residuals

Estimate B0, B1, error

How to fit best fit line
- Average residual = 0
- Minimize sum of squared errors (SSE)    

Yi = (y_i-yHat)*2

SSE+ SSR  = (TSS total sum of squares)

The strength regression has to do with the balance between SSE in relation to TXX 

Assumptions of regression
1. Continuous Y
2. Normally distributed residuals
3. *Linear relationship* *Important*
4. No heteroscedasticity
5. Independent samples (No autocorrelation)


*Formal tests only valuable when data sizes are small
  *With large dataset you're going to get 


-------

T-test is 2 groups


What's an ANOVA?
t-test w more than 2 groups

You get an F-statistic from an ANOVA
- testing the significance of [factor]


Basic idea of interactions:
- The effect of one variable depends on the value of another variable



```{r}
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
```

```{r}
data("mtcars")

mtcars %>% 
ggplot(aes(x=wt, y=mpg)) +
  geom_smooth() +
  geom_point()
```


```{r}
lm1 <- lm(mpg~wt, data = mtcars)
summary(lm1) # p-value 1.29e-10

cor.test(mtcars$wt, mtcars$mpg) # A classical test gives the exact same result

ggplot(lm1, aes(y=.resid, x=.fitted)) +
  geom_point()

plot(lm1)

```




```{r}
bull.rich <- read.csv("./data/Bull_richness.csv")

bull.rich %>%
  filter(GrowthStage == "V8" & Treatment == "Conv.") %>%
  ggplot(aes(x=Fungicide, y = richness)) +
  geom_boxplot()

bull.rich.sub <- bull.rich %>% 
  filter(GrowthStage == "V8" & Treatment == "Conv.")
  # Saaaaame
  
t.test()
summary(lm)
anova(lm)



bull.rich.sub2 <- bull.rich %>% 
  filter(Fungicide == "C", Treatment == "Conv.", Crop == "Corn")

bull.rich.sub2 %>%
  ggplot(aes(x=GrowthStage, y = richness)) +
  geom_boxplot()

lm.growth <- lm(richness ~ GrowthStage, data = bull.rich.sub2)
summary(lm.growth)  
anova(lm.growth) # Would show an F table in a paper
```

```{r}
lsmeans <- emmeans(lm.growth, ~GrowthStage)
results <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE)

emmeans(lm.growth, ~GrowthStage)
cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE)

```


Add interaction

```{r}
bull.rich.sub3 <- bull.rich %>% 
  filter(Treatment == "Conv.", Crop == "Corn")

lm.inter <- lm(richness ~ GrowthStage + Fungicide + GrowthStage:Fungicide, data = bull.rich.sub3)

lm.inter <- lm(richness ~ GrowthStage*Fungicide, data = bull.rich.sub3)
summary(lm.inter)
anova(lm.inter) # This is the ANOVA table that you might report in a paper
```


R Graph Gallery looks nice

```{r}
bull.rich.sub3 %>% 
  ggplot(aes(x = GrowthStage, y = richness, fill = Fungicide)) +
  geom_boxplot()
```



Use emmeans to understand these results better


```{r}
lsmeans <- emmeans(lm.inter, ~Fungicide|GrowthStage) # Effect of Fungicide within growth stage
results <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE)
```

Mixed effects model


```{r}
lme1 <- lmer(richness ~ GrowthStage*Fungicide + (1|Rep), data = bull.rich.sub3)
summary(lme1) # Slightly reduced standard error around our betas
# Shows that random effect can improve power
summary(lm.inter)


# To look at differences in 
lsmeans <- emmeans(lme1, ~Fungicide|GrowthStage) # Effect of Fungicide within growth stage
results <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE)
```



